A Theory for Distributed Database Transactions

### 2019-06-13 2018_Transactions/articleTransactions.tex
improved the definition of isolation

	We define $\leq$ to represent the fact that history $\id{prec}$ has preceded history $\id{prec}'$ in time.
	It is defined by:
\begin{equation}
\label{def:lies in the past}
	\id{prec}\leq\id{prec}'\ \Leftrightarrow\ \id{prec}^*\subseteq\id{prec}'^*
\end{equation}

\section{Traces}
\label{sct:Traces}
	A transaction in a DIS may be perceived as a distributed action,
	which leaves a trace of events in the history of a DIS.
	Let $\mathbb T$ be the set of all conceivable traces.
	With \id{prec} being a history, we define that every pair $(e_1,e_2)\in\id{prec}^*$ is a trace.
	So the traces of history $\id{prec}$ are denoted ${\mathcal T}_{\id{prec}}$ with ${\mathcal T}_{\id{prec}}\subseteq\mathbb T$.
\begin{definition}[Traces]
\label{def:Traces}
\item   The set of traces of history $\text{prec}$ is defined by ${\mathcal T}_{\text{prec}}\ =\ \text{prec}^*$.
\end{definition}
	Pair $(e,e)$ represents the empty trace for any $e\in\id{events}(\id{prec})$ because $(e,e)\in\id{prec}^*$.
	We will denote the empty trace as $0$ and define $(e_1,e_1)=0=(e_2,e_2)$.
	If $(e_1,e)\in\id{prec}^*$ and $(e,e_2)\in\id{prec}'^*$ then
	we define the concatenation operator $+:\mathbb T\rightarrow\mathbb T\rightarrow\mathbb T$ by
\begin{definition}[Concatenation of traces]
\label{def:concatenation}
\item   $(e_1,e)+(e,e_2)=(e_1,e_2)$
\end{definition}
	Note that the definition of $+$ is partial, because $(e_1,e_2)+(e_3,e_4)$ is not defined unless $e_2=e_3$.
	Also, $(e_1,e)+(e,e_2)$ is defined in the combined history $\id{prec}^*\cup\id{prec}'^*$
	because $(e_1,e_2)\in(\id{prec}\cup\id{prec}')^*$.

@@

	It makes sense to define the notion of prefix, so we can define notions like shared history.
\begin{definition}[Prefix]
\label{def:Prefix}
\begin{eqnarray}
	x\leq y&\Rightarrow&\exists z:\ x+z=y
\end{eqnarray}
\end{definition}
	We also define subtraction of traces as a way to "cut off" a prefix from a trace.
\begin{definition}[Subtract]
\label{def:Subtract}
\begin{array}[t]{rcll}
	y-x&=&\text{if }x\leq y:&\exists z: y=x+z\Rightarrow z\\
	   & &\text{otherwise}:&0
\end{array}
\end{definition}

@@

	Isolation is meant to express that the effect of an action depends only on its current state.
\begin{definition}[Isolation]
\item An action $\id{act}:\mathbb H\rightarrow\mathbb H$ works in isolation iff there is a function $f:\mathbb S\rightarrow\mathbb S$ such that:
\label{def:Isolation}
\begin{eqnarray}
	(h',e')=\id{act}(h,e)&\Rightarrow&\id{stat}(h',e')=f(\id{stat}(h,e))
\end{eqnarray}
\end{definition}

	Atomicity expresses that a sequence of multiple actions can be treated as one.
\begin{definition}[Atomicity]
\item An action $\id{act}:\mathbb H\rightarrow\mathbb H$ is atomic iff its constituent actions ...
\label{def:Atomicity}
\begin{eqnarray}
	(h',e')=\id{act}(h,e)&\Rightarrow&e\ \id{pred}^*\ e'
\end{eqnarray}
\end{definition}

>>>>

	We define $\leq$ to represent the fact that history $\id{prec}$ lies in the past of history $\id{prec}'$.
\begin{definition}[$\leq$]
\label{def:lies in the past}
\begin{displaymath}
	\id{prec}\leq\id{prec}'\ \Leftrightarrow\ \id{prec}^*\subseteq\id{prec}'^*\ \wedge\ \id{initial}(\id{prec})\subseteq\id{initial}(\id{prec}')
\end{displaymath}
\end{definition}

\section{Traces}
\label{sct:Traces}
	A transaction in a DIS may be perceived as a distributed action,
	which leaves a trace of events in the history of a DIS.
	Let $\mathbb T$ be the set of all conceivable traces.
	With \id{prec} being a history, we define a trace $(e_1,e_2)_{\id{prec}}$ as follows:
\begin{definition}[Traces]
\label{def:Traces}
\item   $(e_1,e_2)_{\id{prec}}\in\mathbb T\ \Leftrightarrow\ (e_1,e_2)\in\text{prec}^*$
\end{definition}
	We use variables $x$, $y$, and $z$ to represent traces.
	We 
\begin{definition}[Empty trace]
\label{def:Empty trace}
\item   $\forall e: 0_{\id{prec}}\ =\ (e,e)_{\id{prec}}$
\end{definition}
	We denote the empty trace as $0$ and define $(e_1,e_1)=0=(e_2,e_2)$.
	If $(e_1,e)_{\id{prec}}$ and $(e,e_2)_{\id{prec}'}$ are traces then
	we define the concatenation operator $+:\mathbb T\rightarrow\mathbb T\rightarrow\mathbb T$ by
\begin{definition}[Concatenation of traces]
\label{def:concatenation}
\item   $(e_1,e)_{\id{prec}}+(e,e_2)_{\id{prec}'}=(e_1,e_2)_{\id{prec}\cup\id{prec}'}$
\end{definition}
	Note that the definition of $+$ is partial, because $(e_1,e_2)_{\id{prec}}+(e_3,e_4)_{\id{prec}'}$ is not defined unless $e_2=e_3$.
	Also, the result is defined in the combined history $\id{prec}^*\cup\id{prec}'^*$
	because $(e_1,e_2)\in(\id{prec}\cup\id{prec}')^*$. (proof needed)

@@

	The operator $\leq$ prefix, so we can define notions like shared history.
$
(e_1,e_2)\leq(e_3,e_4)
\Leftrightarrow
(\exists\id{prec}:(e_1,e_2)\in\id{prec}^*)\ \wedge\ (\exists\id{prec}':(e_3,e_4)\in\id{prec}'^*)\ \wedge\ \id{prec}\leq\id{prec}'
$
\begin{eqnarray}
	x\leq y&\Rightarrow&\exists z:\ x+z=y\label{def:Prefix}
\end{eqnarray}
	We also define subtraction of traces as a way to "cut off" a prefix from a trace.
\begin{definition}[Subtract]
\label{def:Subtract}
\begin{array}[t]{rcll}
	y-x&=&\text{if }x\leq y:&\exists z: y=x+z\Rightarrow z\\
	   & &\text{otherwise}:&0
\end{array}
\end{definition}

@@

	An action $\id{act}_\id{prec}(e_1,e_2)$ is performed in isolation if it does not affect any other action in the system.
	This means that the only event in trace $(e_1,e_2)_\id{prec}$ that precedes events outside that trace is e_1.
	Likewise, the only event in trace $(e_1,e_2)_\id{prec}$ that is preceded by events outside that trace is e_2.
	To define this, we need an auxiliary function
	$\id{atomize}:\mathbb E\times\mathbb E\times\mathbb H\rightarrow\mathbb H$,
	which replaces a trace $(e_1,e_2)_\id{prec}$ by the pair $(e_1,e_2)$
	in a history $\id{prec}$.
\begin{eqnarray}
	\begin{array}[t]{r@{}l@{}l}
	\multicolumn{3}{l}{\id{atomize}(e_1,e_2,\id{prec})}\\
	   \ =\ \id{prec}-&(&\{ (e_1,e)\in\id{prec}| e_1\ \id{prec}^*\ e\ \wedge\ e\ \id{prec}^*\ e_2\}\ \cup\\
			  & &\{ (e,e_2)\in\id{prec}| e_1\ \id{prec}^*\ e\ \wedge\ e\ \id{prec}^*\ e_2\}\\
			  &)\\
		\cup&&\{ (e_1,e_2) \}\\
	\end{array}
\end{eqnarray}
	We can now define isolation as the property that the traces of the "atomized" history
	are the same as the atomization of all traces in that history.
\begin{definition}[Isolation]
\item An action $\id{act}_{\id{prec}}(e_1,e_2)$ works in isolation iff:
\label{def:Isolation}
\begin{eqnarray}
	(\id{atomize}(e_1,e_2,\id{prec}))^*&=&\id{atomize}(e_1,e_2,\id{prec}^*)
\end{eqnarray}
\end{definition}

	Atomicity expresses that a sequence of multiple actions can be treated as one.

\begin{definition}[Atomic]
\item An action $\id{act}_{\id{prec}}(e_1,e_2)$ is atomic iff
\label{def:Atomicity}
\end{definition}
